---
title: "NBA Project"
author: "Ilya Kukovitskiy"
date: "May 7, 2018"
output: html_document
self_contained: no
---
```{r,results='hide',message=FALSE}
set.seed(1)
library(tree)
library(randomForest)
library(gbm)
library(ggplot2)
```

## Data Preprocessing

#### Reading in and storing dataset from file

```{r}
d <- read.csv("nbasample2GrpPlus0429.csv",header=T, sep=",")
```

#### Removing ax_ and la_ variales, as requested

```{r}
a <- colnames(d)
AX <- grepl("ax_",a) #seeing which values have prefix ax_
AXN <- c() #creating temporary vector recording positions
for(i in 1:length(a)){ #getting rejected values
  if(AX[i]){
    AXN <- c(AXN,i)
  }
}
d <- d[,-AXN] #removing rejected values
a <- colnames(d)
LA <- grepl("la_",a)
LAN <- c() #creating temporary vector recording positions
for(i in 1:length(a)){ #getting rejected values
  if(LA[i]){
    LAN <- c(LAN,i)
  }
}
d <- d[,-LAN] #removing rejected values
a <- colnames(d)
#Cleaning up Environment
rm(AX,AXN,LA,LAN)
```

#### Removing other Variables

```{r}
#closest_arena and closest_team are identical - removing closest_arena

x <- match("closest_arena",a)
#x
d <- d[,-x]
a <- colnames(d)
#match("closest_arena",a) #successfully removed

#tree() can only handle 32 levels on factor predictors

#names(Filter(is.factor, d)) #taking a look at which variables have factors - 

#zip_team and zip_arena have 56 levels - too many to grow trees on - removing both
x <- match(c("zip_team","zip_arena"),a)
#x
d <- d[,-x]
a <- colnames(d)
#match(c("zip_team","zip_arena"),a) #successfully removed

#email_domain has 1468 levels - too many.. -removing
x <- match(c("email_domain"),a)
#x
d <- d[,-x]
a <- colnames(d)
#match(c("email_domain"),a) #successfully removed

#source_date has 3773 levels - too many.. -removing
x <- match(c("source_date"),a)
#x
d <- d[,-x]
a <- colnames(d)
#match(c("source_date"),a) #successfully removed
```

#### Making a training/test set vector

```{r}
train <- sample(1:nrow(d), nrow(d)*(.8))
```

### Removing other response variables besides post_total_tickets

```{r}
#Finding postion of desired response variable
Q <- match('post_total_tickets',a)
a[Q]
#Removing all other response (post_) variables
POSTN_ptt <- POSTN[! POSTN %in% Q]
#POSTN_ptt
a[POSTN_ptt]
dataset.ptt <- d[,-POSTN_ptt]
```
```{r,results='hide',message=FALSE}
attach(dataset.ptt)
```

#### Looking at distribution of test response variable

```{r}
#Creating Violin Plot
p <- ggplot(dataset.ptt[-train,], aes(y=post_total_tickets, 1))
p + geom_violin()
```

#### That's a lot of zeroes

```{r}
#Finding Number of Zeroes
O <- table(post_total_tickets[-train]==0)
#O[2]
```

#### Growing Preliminary Regression Tree

```{r}
#Growing Tree
tree.ptt <- tree(post_total_tickets~., dataset.ptt)
#Plotting Tree
plot(tree.ptt)
text(tree.ptt,pretty=0)
```

#### Estimate MSE on (extremely similar) Trained Tree

```{r}
#Growing Tree
tree.ptt_train <- tree(post_total_tickets~., dataset.ptt, subset=train)
#Plotting Tree
plot(tree.ptt_train)
text(tree.ptt_train,pretty=0)
```

#### Calculating, Plotting, and Recording (Mean Squared) Error

```{r}
#Fitting model on test set
tree.ptt_pred <- predict(tree.ptt_train, dataset.ptt[-train,])
plot(tree.ptt_pred, post_total_tickets[-train])
abline(0,1)
#Residuals plot
plot(post_total_tickets[-train],(tree.ptt_pred-post_total_tickets[-train]))
#MSE
MSE.ptt <- mean((tree.ptt_pred-post_total_tickets[-train])^2)
MSE.ptt
```
```{r,echo=FALSE}
print(paste("Test MSE is",MSE.ptt,". Root MSE is",sqrt(MSE.ptt),", indicating this model leads to test predictions within around",round(sqrt(MSE.ptt),2),"total tickets purchased."))
#fix Compare with distribution of response through box-plot
```
